---
title: "Sentencing mapping"
format: html
execute:
  echo: true
  warning: false
  error: false
---

This Quarto document provides an R workflow that:

-   reads `data/legal_alcohol_policy.csv` as the single input dataset;
-   builds a single working spatial dataset (world + UK subdivisions);
-   marks rows as `covered` directly from the policy dataset;
-   joins policy fields to each mapped country (where available);
-   plots the selected countries directly to the R plotting pane (no file outputs by default).

Notes: - If `rmapshaper` is installed the document will use `rmapshaper::ms_simplify()` for faster, topology-preserving simplification; otherwise it falls back to `sf::st_simplify()`.

```{r setup}
library(sf)
library(ggplot2)
library(dplyr)

# Paths (edit only these)
policy_path <- "data/legal_alcohol_policy.csv"
uk_shp <- "UK/CTRY_DEC_2022_UK_BFC_v2_c.shp"

world_shp_candidates <- c(
  "World_Countries_Generalized.shp",
  "World_Countries_(Generalized)_-573431906301700955/World_Countries_Generalized.shp"
)
world_shp <- world_shp_candidates[file.exists(world_shp_candidates)][1]
if (is.na(world_shp)) {
  stop("Could not find world shapefile. Checked: ", paste(world_shp_candidates, collapse = ", "))
}

# (No caching: the document always rebuilds `map_data` from the raw shapefiles.)

standardise_country <- function(x) {
  out <- tolower(trimws(as.character(x)))
  alias_map <- c(
    "turkey" = "turkiye",
    "czech republic" = "czechia",
    "united states" = "united states of america"
  )
  out <- ifelse(out %in% names(alias_map), alias_map[out], out)
  out
}

```

## Read policy data

```{r read-data}
if (!file.exists(policy_path)) {
  stop("Policy file not found at ", policy_path)
}

policy_df <- read.csv(policy_path, stringsAsFactors = FALSE, check.names = FALSE)
names(policy_df) <- tolower(names(policy_df))

required_cols <- c("country")
missing_cols <- setdiff(required_cols, names(policy_df))
if (length(missing_cols) > 0) {
  stop("Policy file is missing required columns: ", paste(missing_cols, collapse = ", "))
}

policy_df$country_std <- standardise_country(policy_df$country)
policy_df <- policy_df %>% distinct(country_std, .keep_all = TRUE)

if ("covered" %in% names(policy_df)) {
  cvals <- tolower(trimws(as.character(policy_df$covered)))
  policy_df$covered_policy <- cvals %in% c("y", "yes", "true", "1")
} else {
  policy_df$covered_policy <- TRUE
}

# clean character fields for consistent joins and legends
char_cols <- names(policy_df)[vapply(policy_df, is.character, logical(1))]
for (col in char_cols) {
  policy_df[[col]] <- trimws(policy_df[[col]])
  policy_df[[col]][policy_df[[col]] == ""] <- NA
}

policy_country_names <- unique(policy_df$country_std)

cat("Loaded policy rows:", nrow(policy_df), "\n")
cat("Policy countries:", length(policy_country_names), "\n")
```

## Build working spatial dataset

```{r build-working}
# This document now builds the working spatial dataset directly from the
# raw shapefiles. The combined object is called `map_data` and is kept
# deliberately simple: one `name` column, geometry, a `source` flag,
# placeholders for extra attributes, and a logical `covered` column.
# Keep this code as is. 

world0 <- st_read(world_shp, quiet = TRUE)
uk0 <- st_read(uk_shp, quiet = TRUE)

# choose sensible name fields
wname <- intersect(c("COUNTRY","country","NAME","name","NAME_LONG","name_long","admin","ADMIN"), names(world0))[1]
if (is.na(wname)) stop("No name field found in world shapefile")
uname <- if ("CTRY22NM" %in% names(uk0)) "CTRY22NM" else intersect(names(uk0), c("name","NAME","country","COUNTRY"))[1]
if (is.na(uname)) stop("No name field found in UK shapefile")

# keep only name + geometry to preserve simplicity
world <- world0[, wname, drop = FALSE]
names(world)[1] <- "name"
world$name <- standardise_country(as.character(world$name))
world$source <- "world"

uk <- uk0[, uname, drop = FALSE]
names(uk)[1] <- "name"
uk$name <- standardise_country(as.character(uk$name))
uk$source <- "uk"

# ensure same CRS
if (st_crs(world) != st_crs(uk)) uk <- st_transform(uk, st_crs(world))

# remove aggregate United Kingdom from world if present
world <- world[!world$name %in% c("united kingdom","unitedkingdom","united kingdom of great britain and northern ireland"), ]

# combine into a single working dataset named `map_data`
map_data <- rbind(world, uk)

# start from country membership in policy file
map_data$covered <- map_data$name %in% policy_country_names

# join legal/alcohol policy metadata and apply covered policy flag
policy_join <- policy_df
if ("covered" %in% names(policy_join)) {
  policy_join$covered <- NULL
}

map_data <- map_data %>%
  left_join(policy_join, by = c("name" = "country_std"))

map_data$covered <- ifelse(is.na(map_data$covered_policy), map_data$covered, map_data$covered_policy)

simplify_for_plot <- function(x) {
  x_valid <- suppressWarnings(sf::st_make_valid(x))

  if (requireNamespace("rmapshaper", quietly = TRUE)) {
    return(rmapshaper::ms_simplify(x_valid, keep = 0.08, keep_shapes = TRUE))
  }

  crs_orig <- sf::st_crs(x_valid)
  x_work <- if (sf::st_is_longlat(x_valid)) sf::st_transform(x_valid, 3857) else x_valid
  x_simpl <- sf::st_simplify(x_work, dTolerance = 15000, preserveTopology = TRUE)

  if (!is.na(crs_orig) && sf::st_crs(x_simpl) != crs_orig) {
    x_simpl <- sf::st_transform(x_simpl, crs_orig)
  }

  x_simpl
}

map_plot_data <- simplify_for_plot(map_data)

message("Built working dataset: map_data (rows = ", nrow(map_data), ")")
message("Built simplified plotting dataset: map_plot_data (rows = ", nrow(map_plot_data), ")")

```

```{r diagnostic-matching}
# Diagnostic: report any policy countries not matched in shapefile names
# against the shapefile `map_data$name`. Helpful to spot spelling/standardisation issues.
policy_country_names <- sort(unique(policy_country_names))
matched <- intersect(map_data$name, policy_country_names)
unmatched <- setdiff(policy_country_names, map_data$name)
cat("Policy country names:", length(policy_country_names), "\n")
cat("Matched in shapefile :", length(matched), "\n")
cat("Unmatched           :", length(unmatched), "\n")
if (length(unmatched) > 0) {
  print(unmatched)
  # suggest closest matches by edit distance
  suggestions <- sapply(unmatched, function(u) {
    d <- utils::adist(u, map_data$name)
    map_data$name[which.min(d)]
  }, USE.NAMES = FALSE)
  suggestion_df <- data.frame(requested = unmatched, suggestion = suggestions, stringsAsFactors = FALSE)
  print(suggestion_df)
}
```

## Plot selected countries

```{r plot, fig.height=7, fig.width=11}
# remove Antarctica (and similar polar features) which distorts the map
map_data_no_antar <- map_plot_data[!grepl("antar", map_plot_data$name), ]

if (!("covered" %in% names(map_data_no_antar))) {
  map_data_no_antar$covered <- FALSE
}

highlighted_sf <- map_data_no_antar[!is.na(map_data_no_antar$covered) & map_data_no_antar$covered, ]
covered_count <- nrow(highlighted_sf)

# prepare readable labels
map_data_no_antar$label <- tools::toTitleCase(map_data_no_antar$name)

p <- ggplot() +
  geom_sf(data = map_data_no_antar, fill = "#f0f0f0", colour = "#d0d0d0", size = 0.08) +
  geom_sf(data = highlighted_sf, fill = "#4b4b4b", colour = "#000000", size = 0.2, alpha = 0.95) +
  labs(title = "Selected countries",
    subtitle = paste0("Highlighted: ", covered_count, " countries from policy dataset")) +
  coord_sf(expand = FALSE) +
  theme_void() +
  theme(
    plot.title = element_text(size = 18, face = "bold", margin = margin(b = 6)),
    plot.subtitle = element_text(size = 11, margin = margin(b = 8)),
    plot.caption = element_text(size = 8, margin = margin(t = 8)),
    panel.background = element_rect(fill = "#ffffff", colour = NA),
    plot.background = element_rect(fill = "#ffffff", colour = NA),
    legend.position = "none"
  )

# Add labels for covered countries. 
covered_sf <- highlighted_sf
if (nrow(covered_sf) > 0) {
  pts <- st_point_on_surface(covered_sf)
  pts$label <- tools::toTitleCase(covered_sf$name)
  coords <- st_coordinates(pts)
  labdf <- data.frame(X = coords[,1], Y = coords[,2], label = pts$label, stringsAsFactors = FALSE)
  if (requireNamespace("ggrepel", quietly = TRUE)) {
    p <- p + ggrepel::geom_label_repel(data = labdf,
                                       aes(X, Y, label = label),
                                       size = 4.2,
                                       fontface = "bold",
                                       fill = "#ffffffcc",
                                       colour = "#000000",
                                       box.padding = 0.35,
                                       segment.size = 0.25,
                                       segment.colour = "#444444",
                                       min.segment.length = 0,
                                       max.overlaps = 200)
  } else {
    p <- p + geom_label(data = labdf, aes(X, Y, label = label), size = 4.2, fontface = "bold", fill = "#ffffff", colour = "#000000", label.r = unit(0.15, "lines"))
  }
}

print(p)
```

## Individual variable maps

```{r map-helpers, fig.height=7, fig.width=11}
map_data_no_antar <- map_plot_data[!grepl("antar", map_plot_data$name), ]
pattern_pool <- c("stripe", "crosshatch", "circle", "wave", "none")
uk_xlim <- c(-11, 3)
uk_ylim <- c(49, 61)
uk_focus_names <- c("england", "wales", "scotland", "northern ireland")

add_uk_emphasis <- function(p_main, plot_df, selected_df) {
  uk_outline_df <- plot_df[plot_df$name %in% uk_focus_names, ]
  p_main +
    geom_sf(data = uk_outline_df, fill = NA, colour = "#000000", linewidth = 0.45, show.legend = FALSE) +
    theme(legend.position = "bottom")
}

prepare_map_values <- function(df, var_name) {
  vals <- if (var_name %in% names(df)) {
    trimws(as.character(df[[var_name]]))
  } else {
    rep(NA_character_, nrow(df))
  }

  vals_lower <- tolower(vals)
  vals[vals_lower %in% c("yes", "y")] <- "Yes"
  vals[vals_lower %in% c("no", "n")] <- "No"
  vals[vals_lower %in% c("yes (by state)", "yes by state")] <- "Yes (by state)"
  vals[vals_lower %in% c("spirits only", "yes (for spirits only)", "yes for spirits only", "yes - spirits only")] <- "Yes (for spirits only)"

  vals[is.na(vals) | vals == ""] <- "No data"
  vals[vals == "No data"] <- "No"
  factor(vals)
}

make_style_values <- function(levels_vec) {
  levels_vec <- as.character(levels_vec)
  visible_patterns <- c("stripe", "crosshatch", "circle", "wave")
  visible_fills <- c("#d9d9d9", "#bdbdbd", "#969696", "#737373", "#525252", "#252525")

  pattern_values <- stats::setNames(rep("stripe", length(levels_vec)), levels_vec)
  fill_values <- stats::setNames(rep("#d9d9d9", length(levels_vec)), levels_vec)

  other_levels <- setdiff(levels_vec, c("Yes", "Yes (by state)", "No"))
  if (length(other_levels) > 0) {
    pattern_values[other_levels] <- rep(visible_patterns, length.out = length(other_levels))
    fill_values[other_levels] <- rep(visible_fills, length.out = length(other_levels))
  }

  if ("Yes" %in% levels_vec) {
    pattern_values["Yes"] <- "stripe"
    fill_values["Yes"] <- "#bdbdbd"
  }
  if ("Yes (by state)" %in% levels_vec) {
    pattern_values["Yes (by state)"] <- "crosshatch"
    fill_values["Yes (by state)"] <- "#737373"
  }
  if ("Yes (for spirits only)" %in% levels_vec) {
    pattern_values["Yes (for spirits only)"] <- "circle"
    fill_values["Yes (for spirits only)"] <- "#525252"
  }
  if ("No" %in% levels_vec) {
    pattern_values["No"] <- "none"
    fill_values["No"] <- "#000000"
  }

  list(pattern_values = pattern_values, fill_values = fill_values)
}

required_map_vars <- c(
  "legal_system",
  "specialist_substance_use_courts",
  "alcohol_treatment_requirements",
  "alcohol_tagging_requirements",
  "min_purchase_age_on_premise",
  "min_purchase_age_off_premise",
  "bac_limit_general_pop_per_l",
  "bac_limit_young_drivers_per_l",
  "bac_limit_prof_drivers_per_l",
  "random_breath_testing",
  "national_alcohol_strategy_written_adopted",
  "primary_legislation_illegal_alcohol",
  "national_system_monitoring_consumption",
  "national_system_monitoring_health_consequence",
  "national_system_monitoring_social_consequence",
  "national_system_monitoring_policy_response",
  "average_daily_intake_grams_pop_15+_bothsex"
)

missing_map_vars <- setdiff(required_map_vars, names(map_data_no_antar))
if (length(missing_map_vars) > 0) {
  warning(
    "Missing expected map variable(s): ",
    paste(missing_map_vars, collapse = ", "),
    ". These maps will render as 'No' until the columns are added."
  )
}
```

```{r plot-legal-system, fig.height=7, fig.width=11}
plot_df <- map_data_no_antar
plot_df$map_value <- prepare_map_values(plot_df, "legal_system")
selected_df <- plot_df[!is.na(plot_df$covered) & plot_df$covered, ]

if (requireNamespace("ggpattern", quietly = TRUE)) {
  lvls <- levels(selected_df$map_value)
  style_values <- make_style_values(lvls)
  pattern_values <- style_values$pattern_values
  fill_values <- style_values$fill_values
  p_legal <- ggplot() +
    geom_sf(data = plot_df, fill = "white", colour = "#d0d0d0", size = 0.08) +
    ggpattern::geom_sf_pattern(data = selected_df, aes(pattern = map_value, fill = map_value), colour = "#7f7f7f", size = 0.15,
                               pattern_fill = "#1f1f1f", pattern_colour = "#1f1f1f", pattern_density = 0.45,
                               pattern_spacing = 0.025, pattern_key_scale_factor = 0.7) +
    ggpattern::scale_pattern_manual(values = pattern_values, drop = FALSE, name = "Legal System") +
    scale_fill_manual(values = fill_values, drop = FALSE, name = "Legal System")
} else {
  p_legal <- ggplot() +
    geom_sf(data = plot_df, fill = "white", colour = "#d0d0d0", size = 0.08) +
    geom_sf(data = selected_df, aes(fill = map_value), colour = "#7f7f7f", size = 0.15) +
    scale_fill_manual(values = make_style_values(levels(selected_df$map_value))$fill_values, drop = FALSE, name = "Legal System")
}

print(add_uk_emphasis(
  p_legal +
    labs(title = "Legal System") +
    coord_sf(expand = FALSE) +
    theme_void() +
    theme(plot.title = element_text(size = 18, face = "bold", margin = margin(b = 6)), legend.position = "right"),
  plot_df = plot_df,
  selected_df = selected_df
))
```

```{r plot-specialist-courts, fig.height=7, fig.width=11}
plot_df <- map_data_no_antar
plot_df$map_value <- prepare_map_values(plot_df, "specialist_substance_use_courts")
selected_df <- plot_df[!is.na(plot_df$covered) & plot_df$covered, ]

if (requireNamespace("ggpattern", quietly = TRUE)) {
  lvls <- levels(selected_df$map_value)
  style_values <- make_style_values(lvls)
  pattern_values <- style_values$pattern_values
  fill_values <- style_values$fill_values
  p_courts <- ggplot() +
    geom_sf(data = plot_df, fill = "white", colour = "#d0d0d0", size = 0.08) +
    ggpattern::geom_sf_pattern(data = selected_df, aes(pattern = map_value, fill = map_value), colour = "#7f7f7f", size = 0.15,
                               pattern_fill = "#1f1f1f", pattern_colour = "#1f1f1f", pattern_density = 0.45,
                               pattern_spacing = 0.025, pattern_key_scale_factor = 0.7) +
    ggpattern::scale_pattern_manual(values = pattern_values, drop = FALSE, name = "Specialist Courts") +
    scale_fill_manual(values = fill_values, drop = FALSE, name = "Specialist Courts")
} else {
  p_courts <- ggplot() +
    geom_sf(data = plot_df, fill = "white", colour = "#d0d0d0", size = 0.08) +
    geom_sf(data = selected_df, aes(fill = map_value), colour = "#7f7f7f", size = 0.15) +
    scale_fill_manual(values = make_style_values(levels(selected_df$map_value))$fill_values, drop = FALSE, name = "Specialist Courts")
}

print(add_uk_emphasis(
  p_courts +
    labs(title = "Specialist Substance Use Courts") +
    coord_sf(expand = FALSE) +
    theme_void() +
    theme(plot.title = element_text(size = 18, face = "bold", margin = margin(b = 6)), legend.position = "right"),
  plot_df = plot_df,
  selected_df = selected_df
))
```

```{r plot-treatment, fig.height=7, fig.width=11}
plot_df <- map_data_no_antar
plot_df$map_value <- prepare_map_values(plot_df, "alcohol_treatment_requirements")
selected_df <- plot_df[!is.na(plot_df$covered) & plot_df$covered, ]

if (requireNamespace("ggpattern", quietly = TRUE)) {
  lvls <- levels(selected_df$map_value)
  style_values <- make_style_values(lvls)
  pattern_values <- style_values$pattern_values
  fill_values <- style_values$fill_values
  p_treat <- ggplot() +
    geom_sf(data = plot_df, fill = "white", colour = "#d0d0d0", size = 0.08) +
    ggpattern::geom_sf_pattern(data = selected_df, aes(pattern = map_value, fill = map_value), colour = "#7f7f7f", size = 0.15,
                               pattern_fill = "#1f1f1f", pattern_colour = "#1f1f1f", pattern_density = 0.45,
                               pattern_spacing = 0.025, pattern_key_scale_factor = 0.7) +
    ggpattern::scale_pattern_manual(values = pattern_values, drop = FALSE, name = "Treatment Requirements") +
    scale_fill_manual(values = fill_values, drop = FALSE, name = "Treatment Requirements")
} else {
  p_treat <- ggplot() +
    geom_sf(data = plot_df, fill = "white", colour = "#d0d0d0", size = 0.08) +
    geom_sf(data = selected_df, aes(fill = map_value), colour = "#7f7f7f", size = 0.15) +
    scale_fill_manual(values = make_style_values(levels(selected_df$map_value))$fill_values, drop = FALSE, name = "Treatment Requirements")
}

print(add_uk_emphasis(
  p_treat +
    labs(title = "Alcohol Treatment Requirements") +
    coord_sf(expand = FALSE) +
    theme_void() +
    theme(plot.title = element_text(size = 18, face = "bold", margin = margin(b = 6)), legend.position = "right"),
  plot_df = plot_df,
  selected_df = selected_df
))
```

```{r plot-tagging, fig.height=7, fig.width=11}
plot_df <- map_data_no_antar
plot_df$map_value <- prepare_map_values(plot_df, "alcohol_tagging_requirements")
selected_df <- plot_df[!is.na(plot_df$covered) & plot_df$covered, ]

if (requireNamespace("ggpattern", quietly = TRUE)) {
  lvls <- levels(selected_df$map_value)
  style_values <- make_style_values(lvls)
  pattern_values <- style_values$pattern_values
  fill_values <- style_values$fill_values
  p_tag <- ggplot() +
    geom_sf(data = plot_df, fill = "white", colour = "#d0d0d0", size = 0.08) +
    ggpattern::geom_sf_pattern(data = selected_df, aes(pattern = map_value, fill = map_value), colour = "#7f7f7f", size = 0.15,
                               pattern_fill = "#1f1f1f", pattern_colour = "#1f1f1f", pattern_density = 0.45,
                               pattern_spacing = 0.025, pattern_key_scale_factor = 0.7) +
    ggpattern::scale_pattern_manual(values = pattern_values, drop = FALSE, name = "Tagging Requirements") +
    scale_fill_manual(values = fill_values, drop = FALSE, name = "Tagging Requirements")
} else {
  p_tag <- ggplot() +
    geom_sf(data = plot_df, fill = "white", colour = "#d0d0d0", size = 0.08) +
    geom_sf(data = selected_df, aes(fill = map_value), colour = "#7f7f7f", size = 0.15) +
    scale_fill_manual(values = make_style_values(levels(selected_df$map_value))$fill_values, drop = FALSE, name = "Tagging Requirements")
}

print(add_uk_emphasis(
  p_tag +
    labs(title = "Alcohol Tagging Requirements") +
    guides(
      pattern = guide_legend(ncol = 1, byrow = TRUE),
      fill = guide_legend(
        ncol = 1,
        byrow = TRUE,
        label.hjust = 0,
        label.theme = element_text(size = 7, lineheight = 0.9)
      )
    ) +
    coord_sf(expand = FALSE) +
    theme_void() +
    theme(
      plot.title = element_text(size = 18, face = "bold", margin = margin(b = 6)),
      legend.position = "bottom",
      legend.title = element_text(size = 9),
      legend.text = element_text(size = 7)
    ),
  plot_df = plot_df,
  selected_df = selected_df
))
```

```{r plot-min-age-on, fig.height=7, fig.width=11}
plot_df <- map_data_no_antar
plot_df$map_value <- prepare_map_values(plot_df, "min_purchase_age_on_premise")
selected_df <- plot_df[!is.na(plot_df$covered) & plot_df$covered, ]

if (requireNamespace("ggpattern", quietly = TRUE)) {
  lvls <- levels(selected_df$map_value)
  style_values <- make_style_values(lvls)
  pattern_values <- style_values$pattern_values
  fill_values <- style_values$fill_values
  p_min_on <- ggplot() +
    geom_sf(data = plot_df, fill = "white", colour = "#d0d0d0", size = 0.08) +
    ggpattern::geom_sf_pattern(data = selected_df, aes(pattern = map_value, fill = map_value), colour = "#7f7f7f", size = 0.15,
                               pattern_fill = "#1f1f1f", pattern_colour = "#1f1f1f", pattern_density = 0.45,
                               pattern_spacing = 0.025, pattern_key_scale_factor = 0.7) +
    ggpattern::scale_pattern_manual(values = pattern_values, drop = FALSE, name = "Minimum Age On Premise") +
    scale_fill_manual(values = fill_values, drop = FALSE, name = "Minimum Age On Premise")
} else {
  p_min_on <- ggplot() +
    geom_sf(data = plot_df, fill = "white", colour = "#d0d0d0", size = 0.08) +
    geom_sf(data = selected_df, aes(fill = map_value), colour = "#7f7f7f", size = 0.15) +
    scale_fill_manual(values = make_style_values(levels(selected_df$map_value))$fill_values, drop = FALSE, name = "Minimum Age On Premise")
}

print(add_uk_emphasis(
  p_min_on +
    labs(title = "Minimum Purchase Age (On Premise)") +
    coord_sf(expand = FALSE) +
    theme_void() +
    theme(plot.title = element_text(size = 18, face = "bold", margin = margin(b = 6)), legend.position = "right"),
  plot_df = plot_df,
  selected_df = selected_df
))
```

```{r plot-min-age-off, fig.height=7, fig.width=11}
plot_df <- map_data_no_antar
plot_df$map_value <- prepare_map_values(plot_df, "min_purchase_age_off_premise")
selected_df <- plot_df[!is.na(plot_df$covered) & plot_df$covered, ]

if (requireNamespace("ggpattern", quietly = TRUE)) {
  lvls <- levels(selected_df$map_value)
  style_values <- make_style_values(lvls)
  pattern_values <- style_values$pattern_values
  fill_values <- style_values$fill_values
  p_min_off <- ggplot() +
    geom_sf(data = plot_df, fill = "white", colour = "#d0d0d0", size = 0.08) +
    ggpattern::geom_sf_pattern(data = selected_df, aes(pattern = map_value, fill = map_value), colour = "#7f7f7f", size = 0.15,
                               pattern_fill = "#1f1f1f", pattern_colour = "#1f1f1f", pattern_density = 0.45,
                               pattern_spacing = 0.025, pattern_key_scale_factor = 0.7) +
    ggpattern::scale_pattern_manual(values = pattern_values, drop = FALSE, name = "Minimum Age Off Premise") +
    scale_fill_manual(values = fill_values, drop = FALSE, name = "Minimum Age Off Premise")
} else {
  p_min_off <- ggplot() +
    geom_sf(data = plot_df, fill = "white", colour = "#d0d0d0", size = 0.08) +
    geom_sf(data = selected_df, aes(fill = map_value), colour = "#7f7f7f", size = 0.15) +
    scale_fill_manual(values = make_style_values(levels(selected_df$map_value))$fill_values, drop = FALSE, name = "Minimum Age Off Premise")
}

print(add_uk_emphasis(
  p_min_off +
    labs(title = "Minimum Purchase Age (Off Premise)") +
    coord_sf(expand = FALSE) +
    theme_void() +
    theme(plot.title = element_text(size = 18, face = "bold", margin = margin(b = 6)), legend.position = "right"),
  plot_df = plot_df,
  selected_df = selected_df
))
```

```{r plot-bac-general, fig.height=7, fig.width=11}
plot_df <- map_data_no_antar
plot_df$map_value <- prepare_map_values(plot_df, "bac_limit_general_pop_per_l")
selected_df <- plot_df[!is.na(plot_df$covered) & plot_df$covered, ]

if (requireNamespace("ggpattern", quietly = TRUE)) {
  lvls <- levels(selected_df$map_value)
  style_values <- make_style_values(lvls)
  pattern_values <- style_values$pattern_values
  fill_values <- style_values$fill_values
  p_bac_general <- ggplot() +
    geom_sf(data = plot_df, fill = "white", colour = "#d0d0d0", size = 0.08) +
    ggpattern::geom_sf_pattern(data = selected_df, aes(pattern = map_value, fill = map_value), colour = "#7f7f7f", size = 0.15,
                               pattern_fill = "#1f1f1f", pattern_colour = "#1f1f1f", pattern_density = 0.45,
                               pattern_spacing = 0.025, pattern_key_scale_factor = 0.7) +
    ggpattern::scale_pattern_manual(values = pattern_values, drop = FALSE, name = "BAC General") +
    scale_fill_manual(values = fill_values, drop = FALSE, name = "BAC General")
} else {
  p_bac_general <- ggplot() +
    geom_sf(data = plot_df, fill = "white", colour = "#d0d0d0", size = 0.08) +
    geom_sf(data = selected_df, aes(fill = map_value), colour = "#7f7f7f", size = 0.15) +
    scale_fill_manual(values = make_style_values(levels(selected_df$map_value))$fill_values, drop = FALSE, name = "BAC General")
}

print(add_uk_emphasis(
  p_bac_general +
    labs(title = "BAC Limit (General Population)") +
    coord_sf(expand = FALSE) +
    theme_void() +
    theme(plot.title = element_text(size = 18, face = "bold", margin = margin(b = 6)), legend.position = "right"),
  plot_df = plot_df,
  selected_df = selected_df
))
```

```{r plot-bac-young, fig.height=7, fig.width=11}
plot_df <- map_data_no_antar
plot_df$map_value <- prepare_map_values(plot_df, "bac_limit_young_drivers_per_l")
selected_df <- plot_df[!is.na(plot_df$covered) & plot_df$covered, ]

if (requireNamespace("ggpattern", quietly = TRUE)) {
  lvls <- levels(selected_df$map_value)
  style_values <- make_style_values(lvls)
  pattern_values <- style_values$pattern_values
  fill_values <- style_values$fill_values
  p_bac_young <- ggplot() +
    geom_sf(data = plot_df, fill = "white", colour = "#d0d0d0", size = 0.08) +
    ggpattern::geom_sf_pattern(data = selected_df, aes(pattern = map_value, fill = map_value), colour = "#7f7f7f", size = 0.15,
                               pattern_fill = "#1f1f1f", pattern_colour = "#1f1f1f", pattern_density = 0.45,
                               pattern_spacing = 0.025, pattern_key_scale_factor = 0.7) +
    ggpattern::scale_pattern_manual(values = pattern_values, drop = FALSE, name = "BAC Young Drivers") +
    scale_fill_manual(values = fill_values, drop = FALSE, name = "BAC Young Drivers")
} else {
  p_bac_young <- ggplot() +
    geom_sf(data = plot_df, fill = "white", colour = "#d0d0d0", size = 0.08) +
    geom_sf(data = selected_df, aes(fill = map_value), colour = "#7f7f7f", size = 0.15) +
    scale_fill_manual(values = make_style_values(levels(selected_df$map_value))$fill_values, drop = FALSE, name = "BAC Young Drivers")
}

print(add_uk_emphasis(
  p_bac_young +
    labs(title = "BAC Limit (Young Drivers)") +
    coord_sf(expand = FALSE) +
    theme_void() +
    theme(plot.title = element_text(size = 18, face = "bold", margin = margin(b = 6)), legend.position = "right"),
  plot_df = plot_df,
  selected_df = selected_df
))
```

```{r plot-bac-prof, fig.height=7, fig.width=11}
plot_df <- map_data_no_antar
plot_df$map_value <- prepare_map_values(plot_df, "bac_limit_prof_drivers_per_l")
selected_df <- plot_df[!is.na(plot_df$covered) & plot_df$covered, ]

if (requireNamespace("ggpattern", quietly = TRUE)) {
  lvls <- levels(selected_df$map_value)
  style_values <- make_style_values(lvls)
  pattern_values <- style_values$pattern_values
  fill_values <- style_values$fill_values
  p_bac_prof <- ggplot() +
    geom_sf(data = plot_df, fill = "white", colour = "#d0d0d0", size = 0.08) +
    ggpattern::geom_sf_pattern(data = selected_df, aes(pattern = map_value, fill = map_value), colour = "#7f7f7f", size = 0.15,
                               pattern_fill = "#1f1f1f", pattern_colour = "#1f1f1f", pattern_density = 0.45,
                               pattern_spacing = 0.025, pattern_key_scale_factor = 0.7) +
    ggpattern::scale_pattern_manual(values = pattern_values, drop = FALSE, name = "BAC Professional") +
    scale_fill_manual(values = fill_values, drop = FALSE, name = "BAC Professional")
} else {
  p_bac_prof <- ggplot() +
    geom_sf(data = plot_df, fill = "white", colour = "#d0d0d0", size = 0.08) +
    geom_sf(data = selected_df, aes(fill = map_value), colour = "#7f7f7f", size = 0.15) +
    scale_fill_manual(values = make_style_values(levels(selected_df$map_value))$fill_values, drop = FALSE, name = "BAC Professional")
}

print(add_uk_emphasis(
  p_bac_prof +
    labs(title = "BAC Limit (Professional Drivers)") +
    coord_sf(expand = FALSE) +
    theme_void() +
    theme(plot.title = element_text(size = 18, face = "bold", margin = margin(b = 6)), legend.position = "right"),
  plot_df = plot_df,
  selected_df = selected_df
))
```

```{r plot-breath-testing, fig.height=7, fig.width=11}
plot_df <- map_data_no_antar
plot_df$map_value <- prepare_map_values(plot_df, "random_breath_testing")
selected_df <- plot_df[!is.na(plot_df$covered) & plot_df$covered, ]

if (requireNamespace("ggpattern", quietly = TRUE)) {
  lvls <- levels(selected_df$map_value)
  style_values <- make_style_values(lvls)
  pattern_values <- style_values$pattern_values
  fill_values <- style_values$fill_values
  p_rbt <- ggplot() +
    geom_sf(data = plot_df, fill = "white", colour = "#d0d0d0", size = 0.08) +
    ggpattern::geom_sf_pattern(data = selected_df, aes(pattern = map_value, fill = map_value), colour = "#7f7f7f", size = 0.15,
                               pattern_fill = "#1f1f1f", pattern_colour = "#1f1f1f", pattern_density = 0.45,
                               pattern_spacing = 0.025, pattern_key_scale_factor = 0.7) +
    ggpattern::scale_pattern_manual(values = pattern_values, drop = FALSE, name = "Random Breath Testing") +
    scale_fill_manual(values = fill_values, drop = FALSE, name = "Random Breath Testing")
} else {
  p_rbt <- ggplot() +
    geom_sf(data = plot_df, fill = "white", colour = "#d0d0d0", size = 0.08) +
    geom_sf(data = selected_df, aes(fill = map_value), colour = "#7f7f7f", size = 0.15) +
    scale_fill_manual(values = make_style_values(levels(selected_df$map_value))$fill_values, drop = FALSE, name = "Random Breath Testing")
}

print(add_uk_emphasis(
  p_rbt +
    labs(title = "Random Breath Testing") +
    coord_sf(expand = FALSE) +
    theme_void() +
    theme(plot.title = element_text(size = 18, face = "bold", margin = margin(b = 6)), legend.position = "right"),
  plot_df = plot_df,
  selected_df = selected_df
))
```

```{r plot-national-strategy, fig.height=7, fig.width=11}
plot_df <- map_data_no_antar
plot_df$map_value <- prepare_map_values(plot_df, "national_alcohol_strategy_written_adopted")
selected_df <- plot_df[!is.na(plot_df$covered) & plot_df$covered, ]

if (requireNamespace("ggpattern", quietly = TRUE)) {
  lvls <- levels(selected_df$map_value)
  style_values <- make_style_values(lvls)
  pattern_values <- style_values$pattern_values
  fill_values <- style_values$fill_values
  p_strategy <- ggplot() +
    geom_sf(data = plot_df, fill = "white", colour = "#d0d0d0", size = 0.08) +
    ggpattern::geom_sf_pattern(data = selected_df, aes(pattern = map_value, fill = map_value), colour = "#7f7f7f", size = 0.15,
                               pattern_fill = "#1f1f1f", pattern_colour = "#1f1f1f", pattern_density = 0.45,
                               pattern_spacing = 0.025, pattern_key_scale_factor = 0.7) +
    ggpattern::scale_pattern_manual(values = pattern_values, drop = FALSE, name = "National Strategy") +
    scale_fill_manual(values = fill_values, drop = FALSE, name = "National Strategy")
} else {
  p_strategy <- ggplot() +
    geom_sf(data = plot_df, fill = "white", colour = "#d0d0d0", size = 0.08) +
    geom_sf(data = selected_df, aes(fill = map_value), colour = "#7f7f7f", size = 0.15) +
    scale_fill_manual(values = make_style_values(levels(selected_df$map_value))$fill_values, drop = FALSE, name = "National Strategy")
}

print(add_uk_emphasis(
  p_strategy +
    labs(title = "National Alcohol Strategy (Written & Adopted)") +
    coord_sf(expand = FALSE) +
    theme_void() +
    theme(plot.title = element_text(size = 18, face = "bold", margin = margin(b = 6)), legend.position = "right"),
  plot_df = plot_df,
  selected_df = selected_df
))
```

```{r plot-illegal-alcohol-legislation, fig.height=7, fig.width=11}
plot_df <- map_data_no_antar
plot_df$map_value <- prepare_map_values(plot_df, "primary_legislation_illegal_alcohol")
selected_df <- plot_df[!is.na(plot_df$covered) & plot_df$covered, ]

if (requireNamespace("ggpattern", quietly = TRUE)) {
  lvls <- levels(selected_df$map_value)
  style_values <- make_style_values(lvls)
  pattern_values <- style_values$pattern_values
  fill_values <- style_values$fill_values
  p_illegal <- ggplot() +
    geom_sf(data = plot_df, fill = "white", colour = "#d0d0d0", size = 0.08) +
    ggpattern::geom_sf_pattern(data = selected_df, aes(pattern = map_value, fill = map_value), colour = "#7f7f7f", size = 0.15,
                               pattern_fill = "#1f1f1f", pattern_colour = "#1f1f1f", pattern_density = 0.45,
                               pattern_spacing = 0.025, pattern_key_scale_factor = 0.7) +
    ggpattern::scale_pattern_manual(values = pattern_values, drop = FALSE, name = "Illegal Alcohol Legislation") +
    scale_fill_manual(values = fill_values, drop = FALSE, name = "Illegal Alcohol Legislation")
} else {
  p_illegal <- ggplot() +
    geom_sf(data = plot_df, fill = "white", colour = "#d0d0d0", size = 0.08) +
    geom_sf(data = selected_df, aes(fill = map_value), colour = "#7f7f7f", size = 0.15) +
    scale_fill_manual(values = make_style_values(levels(selected_df$map_value))$fill_values, drop = FALSE, name = "Illegal Alcohol Legislation")
}

print(add_uk_emphasis(
  p_illegal +
    labs(title = "Primary Legislation for Illegal Alcohol") +
    coord_sf(expand = FALSE) +
    theme_void() +
    theme(plot.title = element_text(size = 18, face = "bold", margin = margin(b = 6)), legend.position = "right"),
  plot_df = plot_df,
  selected_df = selected_df
))
```

```{r plot-monitoring-consumption, fig.height=7, fig.width=11}
plot_df <- map_data_no_antar
plot_df$map_value <- prepare_map_values(plot_df, "national_system_monitoring_consumption")
selected_df <- plot_df[!is.na(plot_df$covered) & plot_df$covered, ]

if (requireNamespace("ggpattern", quietly = TRUE)) {
  lvls <- levels(selected_df$map_value)
  style_values <- make_style_values(lvls)
  pattern_values <- style_values$pattern_values
  fill_values <- style_values$fill_values
  p_mon_cons <- ggplot() +
    geom_sf(data = plot_df, fill = "white", colour = "#d0d0d0", size = 0.08) +
    ggpattern::geom_sf_pattern(data = selected_df, aes(pattern = map_value, fill = map_value), colour = "#7f7f7f", size = 0.15,
                               pattern_fill = "#1f1f1f", pattern_colour = "#1f1f1f", pattern_density = 0.45,
                               pattern_spacing = 0.025, pattern_key_scale_factor = 0.7) +
    ggpattern::scale_pattern_manual(values = pattern_values, drop = FALSE, name = "Monitoring Consumption") +
    scale_fill_manual(values = fill_values, drop = FALSE, name = "Monitoring Consumption")
} else {
  p_mon_cons <- ggplot() +
    geom_sf(data = plot_df, fill = "white", colour = "#d0d0d0", size = 0.08) +
    geom_sf(data = selected_df, aes(fill = map_value), colour = "#7f7f7f", size = 0.15) +
    scale_fill_manual(values = make_style_values(levels(selected_df$map_value))$fill_values, drop = FALSE, name = "Monitoring Consumption")
}

print(add_uk_emphasis(
  p_mon_cons +
    labs(title = "Monitoring System: Consumption") +
    coord_sf(expand = FALSE) +
    theme_void() +
    theme(plot.title = element_text(size = 18, face = "bold", margin = margin(b = 6)), legend.position = "right"),
  plot_df = plot_df,
  selected_df = selected_df
))
```

```{r plot-monitoring-health, fig.height=7, fig.width=11}
plot_df <- map_data_no_antar
plot_df$map_value <- prepare_map_values(plot_df, "national_system_monitoring_health_consequence")
selected_df <- plot_df[!is.na(plot_df$covered) & plot_df$covered, ]

if (requireNamespace("ggpattern", quietly = TRUE)) {
  lvls <- levels(selected_df$map_value)
  style_values <- make_style_values(lvls)
  pattern_values <- style_values$pattern_values
  fill_values <- style_values$fill_values
  p_mon_health <- ggplot() +
    geom_sf(data = plot_df, fill = "white", colour = "#d0d0d0", size = 0.08) +
    ggpattern::geom_sf_pattern(data = selected_df, aes(pattern = map_value, fill = map_value), colour = "#7f7f7f", size = 0.15,
                               pattern_fill = "#1f1f1f", pattern_colour = "#1f1f1f", pattern_density = 0.45,
                               pattern_spacing = 0.025, pattern_key_scale_factor = 0.7) +
    ggpattern::scale_pattern_manual(values = pattern_values, drop = FALSE, name = "Monitoring Health") +
    scale_fill_manual(values = fill_values, drop = FALSE, name = "Monitoring Health")
} else {
  p_mon_health <- ggplot() +
    geom_sf(data = plot_df, fill = "white", colour = "#d0d0d0", size = 0.08) +
    geom_sf(data = selected_df, aes(fill = map_value), colour = "#7f7f7f", size = 0.15) +
    scale_fill_manual(values = make_style_values(levels(selected_df$map_value))$fill_values, drop = FALSE, name = "Monitoring Health")
}

print(add_uk_emphasis(
  p_mon_health +
    labs(title = "Monitoring System: Health Consequences") +
    coord_sf(expand = FALSE) +
    theme_void() +
    theme(plot.title = element_text(size = 18, face = "bold", margin = margin(b = 6)), legend.position = "right"),
  plot_df = plot_df,
  selected_df = selected_df
))
```

```{r plot-monitoring-social, fig.height=7, fig.width=11}
plot_df <- map_data_no_antar
plot_df$map_value <- prepare_map_values(plot_df, "national_system_monitoring_social_consequence")
selected_df <- plot_df[!is.na(plot_df$covered) & plot_df$covered, ]

if (requireNamespace("ggpattern", quietly = TRUE)) {
  lvls <- levels(selected_df$map_value)
  style_values <- make_style_values(lvls)
  pattern_values <- style_values$pattern_values
  fill_values <- style_values$fill_values
  p_mon_social <- ggplot() +
    geom_sf(data = plot_df, fill = "white", colour = "#d0d0d0", size = 0.08) +
    ggpattern::geom_sf_pattern(data = selected_df, aes(pattern = map_value, fill = map_value), colour = "#7f7f7f", size = 0.15,
                               pattern_fill = "#1f1f1f", pattern_colour = "#1f1f1f", pattern_density = 0.45,
                               pattern_spacing = 0.025, pattern_key_scale_factor = 0.7) +
    ggpattern::scale_pattern_manual(values = pattern_values, drop = FALSE, name = "Monitoring Social") +
    scale_fill_manual(values = fill_values, drop = FALSE, name = "Monitoring Social")
} else {
  p_mon_social <- ggplot() +
    geom_sf(data = plot_df, fill = "white", colour = "#d0d0d0", size = 0.08) +
    geom_sf(data = selected_df, aes(fill = map_value), colour = "#7f7f7f", size = 0.15) +
    scale_fill_manual(values = make_style_values(levels(selected_df$map_value))$fill_values, drop = FALSE, name = "Monitoring Social")
}

print(add_uk_emphasis(
  p_mon_social +
    labs(title = "Monitoring System: Social Consequences") +
    coord_sf(expand = FALSE) +
    theme_void() +
    theme(plot.title = element_text(size = 18, face = "bold", margin = margin(b = 6)), legend.position = "right"),
  plot_df = plot_df,
  selected_df = selected_df
))
```

```{r plot-monitoring-policy, fig.height=7, fig.width=11}
plot_df <- map_data_no_antar
plot_df$map_value <- prepare_map_values(plot_df, "national_system_monitoring_policy_response")
selected_df <- plot_df[!is.na(plot_df$covered) & plot_df$covered, ]

if (requireNamespace("ggpattern", quietly = TRUE)) {
  lvls <- levels(selected_df$map_value)
  style_values <- make_style_values(lvls)
  pattern_values <- style_values$pattern_values
  fill_values <- style_values$fill_values
  p_mon_policy <- ggplot() +
    geom_sf(data = plot_df, fill = "white", colour = "#d0d0d0", size = 0.08) +
    ggpattern::geom_sf_pattern(data = selected_df, aes(pattern = map_value, fill = map_value), colour = "#7f7f7f", size = 0.15,
                               pattern_fill = "#1f1f1f", pattern_colour = "#1f1f1f", pattern_density = 0.45,
                               pattern_spacing = 0.025, pattern_key_scale_factor = 0.7) +
    ggpattern::scale_pattern_manual(values = pattern_values, drop = FALSE, name = "Monitoring Policy") +
    scale_fill_manual(values = fill_values, drop = FALSE, name = "Monitoring Policy")
} else {
  p_mon_policy <- ggplot() +
    geom_sf(data = plot_df, fill = "white", colour = "#d0d0d0", size = 0.08) +
    geom_sf(data = selected_df, aes(fill = map_value), colour = "#7f7f7f", size = 0.15) +
    scale_fill_manual(values = make_style_values(levels(selected_df$map_value))$fill_values, drop = FALSE, name = "Monitoring Policy")
}

print(add_uk_emphasis(
  p_mon_policy +
    labs(title = "Monitoring System: Policy Response") +
    coord_sf(expand = FALSE) +
    theme_void() +
    theme(plot.title = element_text(size = 18, face = "bold", margin = margin(b = 6)), legend.position = "right"),
  plot_df = plot_df,
  selected_df = selected_df
))
```

```{r plot-average-intake, fig.height=7, fig.width=11}
plot_df <- map_data_no_antar
plot_df$map_value <- prepare_map_values(plot_df, "average_daily_intake_grams_pop_15+_bothsex")
selected_df <- plot_df[!is.na(plot_df$covered) & plot_df$covered, ]

if (requireNamespace("ggpattern", quietly = TRUE)) {
  lvls <- levels(selected_df$map_value)
  style_values <- make_style_values(lvls)
  pattern_values <- style_values$pattern_values
  fill_values <- style_values$fill_values
  p_intake <- ggplot() +
    geom_sf(data = plot_df, fill = "white", colour = "#d0d0d0", size = 0.08) +
    ggpattern::geom_sf_pattern(data = selected_df, aes(pattern = map_value, fill = map_value), colour = "#7f7f7f", size = 0.15,
                               pattern_fill = "#1f1f1f", pattern_colour = "#1f1f1f", pattern_density = 0.45,
                               pattern_spacing = 0.025, pattern_key_scale_factor = 0.7) +
    ggpattern::scale_pattern_manual(values = pattern_values, drop = FALSE, name = "Average Intake") +
    scale_fill_manual(values = fill_values, drop = FALSE, name = "Average Intake")
} else {
  p_intake <- ggplot() +
    geom_sf(data = plot_df, fill = "white", colour = "#d0d0d0", size = 0.08) +
    geom_sf(data = selected_df, aes(fill = map_value), colour = "#7f7f7f", size = 0.15) +
    scale_fill_manual(values = make_style_values(levels(selected_df$map_value))$fill_values, drop = FALSE, name = "Average Intake")
}

print(add_uk_emphasis(
  p_intake +
    labs(title = "Average Daily Intake (grams, 15+, both sexes)") +
    coord_sf(expand = FALSE) +
    theme_void() +
    theme(plot.title = element_text(size = 18, face = "bold", margin = margin(b = 6)), legend.position = "right"),
  plot_df = plot_df,
  selected_df = selected_df
))
```

## Quick edit: add or edit variables

You can edit variables directly in `data/legal_alcohol_policy.csv` and rerun the document. To add a new map, copy one of the chunks above and change the variable name/title. Example in-session edits:

```{r edit-example}
# Example: temporary in-session updates before plotting
map_data$legal_system[map_data$name %in% c("singapore")] <- "common law"
map_data$specialist_substance_use_courts[map_data$name %in% c("singapore")] <- "No"
```
