---
title: "Sentencing mapping"
format: html
execute:
  echo: true
  warning: false
  error: false
---

This Quarto document provides an R workflow that:

-   reads `data/legal_alcohol_policy.csv` as the single input dataset;
-   builds a single working spatial dataset (world + UK subdivisions);
-   marks rows as `covered` directly from the policy dataset;
-   joins policy fields to each mapped country (where available);
-   plots the selected countries directly to the R plotting pane (no file outputs by default).

Notes: - If `rmapshaper` is installed the document will use `rmapshaper::ms_simplify()` for faster, topology-preserving simplification; otherwise it falls back to `sf::st_simplify()`.

```{r setup}
library(sf)
library(ggplot2)
library(dplyr)

# Paths (edit only these)
policy_path <- "data/legal_alcohol_policy.csv"
uk_shp <- "UK/CTRY_DEC_2022_UK_BFC_v2_c.shp"

world_shp_candidates <- c(
  "World_Countries_Generalized.shp",
  "World_Countries_(Generalized)_-573431906301700955/World_Countries_Generalized.shp"
)
world_shp <- world_shp_candidates[file.exists(world_shp_candidates)][1]
if (is.na(world_shp)) {
  stop("Could not find world shapefile. Checked: ", paste(world_shp_candidates, collapse = ", "))
}

# (No caching: the document always rebuilds `map_data` from the raw shapefiles.)

standardise_country <- function(x) {
  out <- tolower(trimws(as.character(x)))
  alias_map <- c(
    "turkey" = "turkiye",
    "czech republic" = "czechia",
    "united states" = "united states of america"
  )
  out <- ifelse(out %in% names(alias_map), alias_map[out], out)
  out
}

```

## Read policy data

```{r read-data}
if (!file.exists(policy_path)) {
  stop("Policy file not found at ", policy_path)
}

policy_df <- read.csv(policy_path, stringsAsFactors = FALSE)
names(policy_df) <- tolower(names(policy_df))

required_cols <- c("country")
missing_cols <- setdiff(required_cols, names(policy_df))
if (length(missing_cols) > 0) {
  stop("Policy file is missing required columns: ", paste(missing_cols, collapse = ", "))
}

policy_df$country_std <- standardise_country(policy_df$country)
policy_df <- policy_df %>% distinct(country_std, .keep_all = TRUE)

if ("covered" %in% names(policy_df)) {
  cvals <- tolower(trimws(as.character(policy_df$covered)))
  policy_df$covered_policy <- cvals %in% c("y", "yes", "true", "1")
} else {
  policy_df$covered_policy <- TRUE
}

policy_country_names <- unique(policy_df$country_std)

cat("Loaded policy rows:", nrow(policy_df), "\n")
cat("Policy countries:", length(policy_country_names), "\n")
```

## Build working spatial dataset

```{r build-working}
# This document now builds the working spatial dataset directly from the
# raw shapefiles. The combined object is called `map_data` and is kept
# deliberately simple: one `name` column, geometry, a `source` flag,
# placeholders for extra attributes, and a logical `covered` column.
# Keep this code as is. 

world0 <- st_read(world_shp, quiet = TRUE)
uk0 <- st_read(uk_shp, quiet = TRUE)

# choose sensible name fields
wname <- intersect(c("COUNTRY","country","NAME","name","NAME_LONG","name_long","admin","ADMIN"), names(world0))[1]
if (is.na(wname)) stop("No name field found in world shapefile")
uname <- if ("CTRY22NM" %in% names(uk0)) "CTRY22NM" else intersect(names(uk0), c("name","NAME","country","COUNTRY"))[1]
if (is.na(uname)) stop("No name field found in UK shapefile")

# keep only name + geometry to preserve simplicity
world <- world0[, wname, drop = FALSE]
names(world)[1] <- "name"
world$name <- standardise_country(as.character(world$name))
world$source <- "world"

uk <- uk0[, uname, drop = FALSE]
names(uk)[1] <- "name"
uk$name <- standardise_country(as.character(uk$name))
uk$source <- "uk"

# ensure same CRS
if (st_crs(world) != st_crs(uk)) uk <- st_transform(uk, st_crs(world))

# remove aggregate United Kingdom from world if present
world <- world[!world$name %in% c("united kingdom","unitedkingdom","united kingdom of great britain and northern ireland"), ]

# combine into a single working dataset named `map_data`
map_data <- rbind(world, uk)

# start from country membership in policy file
map_data$covered <- map_data$name %in% policy_country_names

# join legal/alcohol policy metadata and apply covered policy flag
policy_join <- policy_df
if ("covered" %in% names(policy_join)) {
  policy_join$covered <- NULL
}

map_data <- map_data %>%
  left_join(policy_join, by = c("name" = "country_std"))

map_data$covered <- ifelse(is.na(map_data$covered_policy), map_data$covered, map_data$covered_policy)

message("Built working dataset: map_data (rows = ", nrow(map_data), ")")

```

```{r diagnostic-matching}
# Diagnostic: report any policy countries not matched in shapefile names
# against the shapefile `map_data$name`. Helpful to spot spelling/standardisation issues.
policy_country_names <- sort(unique(policy_country_names))
matched <- intersect(map_data$name, policy_country_names)
unmatched <- setdiff(policy_country_names, map_data$name)
cat("Policy country names:", length(policy_country_names), "\n")
cat("Matched in shapefile :", length(matched), "\n")
cat("Unmatched           :", length(unmatched), "\n")
if (length(unmatched) > 0) {
  print(unmatched)
  # suggest closest matches by edit distance
  suggestions <- sapply(unmatched, function(u) {
    d <- utils::adist(u, map_data$name)
    map_data$name[which.min(d)]
  }, USE.NAMES = FALSE)
  suggestion_df <- data.frame(requested = unmatched, suggestion = suggestions, stringsAsFactors = FALSE)
  print(suggestion_df)
}
```

## Plot selected countries

```{r plot, fig.height=7, fig.width=11}
# remove Antarctica (and similar polar features) which distorts the map
map_data_no_antar <- map_data[!grepl("antar", map_data$name), ]

if (!("covered" %in% names(map_data_no_antar))) {
  map_data_no_antar$covered <- FALSE
}

highlighted_sf <- map_data_no_antar[!is.na(map_data_no_antar$covered) & map_data_no_antar$covered, ]
covered_count <- nrow(highlighted_sf)

# prepare readable labels
map_data_no_antar$label <- tools::toTitleCase(map_data_no_antar$name)

p <- ggplot() +
  geom_sf(data = map_data_no_antar, fill = "#f0f0f0", colour = "#d0d0d0", size = 0.08) +
  geom_sf(data = highlighted_sf, fill = "#4b4b4b", colour = "#000000", size = 0.2, alpha = 0.95) +
  labs(title = "Selected countries",
    subtitle = paste0("Highlighted: ", covered_count, " countries from policy dataset")) +
  coord_sf(expand = FALSE) +
  theme_void() +
  theme(
    plot.title = element_text(size = 18, face = "bold", margin = margin(b = 6)),
    plot.subtitle = element_text(size = 11, margin = margin(b = 8)),
    plot.caption = element_text(size = 8, margin = margin(t = 8)),
    panel.background = element_rect(fill = "#ffffff", colour = NA),
    plot.background = element_rect(fill = "#ffffff", colour = NA),
    legend.position = "none"
  )

# Add labels for covered countries. 
covered_sf <- highlighted_sf
if (nrow(covered_sf) > 0) {
  pts <- st_point_on_surface(covered_sf)
  pts$label <- tools::toTitleCase(covered_sf$name)
  coords <- st_coordinates(pts)
  labdf <- data.frame(X = coords[,1], Y = coords[,2], label = pts$label, stringsAsFactors = FALSE)
  if (requireNamespace("ggrepel", quietly = TRUE)) {
    p <- p + ggrepel::geom_label_repel(data = labdf,
                                       aes(X, Y, label = label),
                                       size = 4.2,
                                       fontface = "bold",
                                       fill = "#ffffffcc",
                                       colour = "#000000",
                                       box.padding = 0.35,
                                       segment.size = 0.25,
                                       segment.colour = "#444444",
                                       min.segment.length = 0,
                                       max.overlaps = 200)
  } else {
    p <- p + geom_label(data = labdf, aes(X, Y, label = label), size = 4.2, fontface = "bold", fill = "#ffffff", colour = "#000000", label.r = unit(0.15, "lines"))
  }
}

print(p)
```

## Plot legal system families

```{r plot-legal-system, fig.height=7, fig.width=11}
map_data_no_antar <- map_data[!grepl("antar", map_data$name), ]

  p_legal <- ggplot(map_data_no_antar) +
    geom_sf(aes(fill = legal_system), colour = "#d0d0d0", size = 0.08) +
    scale_fill_grey(start = 0.9, end = 0.4, na.value = "#f7f7f7") +
    labs(
      title = "Legal System",
      fill = "Legal System Family"
    ) +
    coord_sf(expand = FALSE) +
    theme_void() +
    theme(
      plot.title = element_text(size = 18, face = "bold", margin = margin(b = 6)),
      plot.subtitle = element_text(size = 11, margin = margin(b = 8)),
      legend.position = "right"
    )

  print(p_legal)
  
  # a blank one as need to find out singapore 

```

## Plot alcohol policy status

```{r plot-alcohol-policy, fig.height=7, fig.width=11}

p_policy <- ggplot(map_data_no_antar) +
  geom_sf(aes(fill = alcohol_policy_status), colour = "#d0d0d0", size = 0.08) +
  scale_fill_grey(start = 0.9, end = 0.4, na.value = "#f7f7f7") +
  labs(
      title = "Alcohol Policy Status",
      subtitle = "National alcohol legislation/policy coding",
      fill = "Alcohol Policy Status"
    ) +
    coord_sf(expand = FALSE) +
    theme_void() +
    theme(
      plot.title = element_text(size = 18, face = "bold", margin = margin(b = 6)),
      plot.subtitle = element_text(size = 11, margin = margin(b = 8)),
      legend.position = "right"
    )
```

## Plot other legal/alcohol indicators

```{r plot-min-age-on, fig.height=7, fig.width=11}
p_min_on <- ggplot(map_data_no_antar) +
  geom_sf(aes(fill = min_purchase_age_on_premise), colour = "#d0d0d0", size = 0.08) +
  scale_fill_grey(start = 0.9, end = 0.4, na.value = "#f7f7f7") +
  labs(title = "Minimum Purchase Age (On Premise)", 
       fill = "Minimum Age On Premise") +
  coord_sf(expand = FALSE) +
  theme_void() +
  theme(plot.title = element_text(size = 18, face = "bold", margin = margin(b = 6)),
        plot.subtitle = element_text(size = 11, margin = margin(b = 8)),
        legend.position = "right")

  print(p_min_on)

```

```{r plot-min-age-off, fig.height=7, fig.width=11}

  p_min_off <- ggplot(map_data_no_antar) +
    geom_sf(aes(fill = min_purchase_age_off_premise), colour = "#d0d0d0", size = 0.08) +
    scale_fill_grey(start = 0.9, end = 0.4, na.value = "#f7f7f7") +
    labs(title = "Minimum Purchase Age (Off Premise)", fill = "Minimum Age Off Premise") +
    coord_sf(expand = FALSE) +
    theme_void() +
    theme(plot.title = element_text(size = 18, face = "bold", margin = margin(b = 6)),
          plot.subtitle = element_text(size = 11, margin = margin(b = 8)),
          legend.position = "right")

  print(p_min_off)
  
```

```{r plot-bac-general, fig.height=7, fig.width=11}

  p_bac_general <- ggplot(map_data_no_antar) +
    geom_sf(aes(fill = bac_limit_general_pop_per_l), colour = "#d0d0d0", size = 0.08) +
    scale_fill_grey(start = 0.9, end = 0.4, na.value = "#f7f7f7") +
    labs(title = "BAC Limit (General Population)", subtitle = "Indicator: bac_limit_general_pop_per_l", fill = "BAC General") +
    coord_sf(expand = FALSE) +
    theme_void() +
    theme(plot.title = element_text(size = 18, face = "bold", margin = margin(b = 6)),
          plot.subtitle = element_text(size = 11, margin = margin(b = 8)),
          legend.position = "right")

  print(p_bac_general)
  
```

```{r plot-bac-young, fig.height=7, fig.width=11}

  p_bac_young <- ggplot(map_data_no_antar) +
    geom_sf(aes(fill = bac_limit_young_drivers_per_l), colour = "#d0d0d0", size = 0.08) +
    scale_fill_grey(start = 0.9, end = 0.4, na.value = "#f7f7f7") +
    labs(title = "BAC Limit (Young Drivers)", subtitle = "Indicator: bac_limit_young_drivers_per_l", fill = "BAC Young Drivers") +
    coord_sf(expand = FALSE) +
    theme_void() +
    theme(plot.title = element_text(size = 18, face = "bold", margin = margin(b = 6)),
          plot.subtitle = element_text(size = 11, margin = margin(b = 8)),
          legend.position = "right")

  print(p_bac_young)
  
```

```{r plot-bac-prof, fig.height=7, fig.width=11}

  p_bac_prof <- ggplot(map_data_no_antar) +
    geom_sf(aes(fill = bac_limit_prof_drivers_per_l), colour = "#d0d0d0", size = 0.08) +
    scale_fill_grey(start = 0.9, end = 0.4, na.value = "#f7f7f7") +
    labs(title = "BAC Limit (Professional Drivers)", subtitle = "Indicator: bac_limit_prof_drivers_per_l", fill = "BAC Professional") +
    coord_sf(expand = FALSE) +
    theme_void() +
    theme(plot.title = element_text(size = 18, face = "bold", margin = margin(b = 6)),
          plot.subtitle = element_text(size = 11, margin = margin(b = 8)),
          legend.position = "right")

  print(p_bac_prof)

```

```{r plot-breath-testing, fig.height=7, fig.width=11}

  p_rbt <- ggplot(map_data_no_antar) +
    geom_sf(aes(fill = random_breath_testing), colour = "#d0d0d0", size = 0.08) +
    scale_fill_grey(start = 0.9, end = 0.4, na.value = "#f7f7f7") +
    labs(title = "Random Breath Testing", subtitle = "Indicator: random_breath_testing", fill = "Random Breath Testing") +
    coord_sf(expand = FALSE) +
    theme_void() +
    theme(plot.title = element_text(size = 18, face = "bold", margin = margin(b = 6)),
          plot.subtitle = element_text(size = 11, margin = margin(b = 8)),
          legend.position = "right")

  print(p_rbt)

```

```{r plot-national-strategy, fig.height=7, fig.width=11}

  p_strategy <- ggplot(map_data_no_antar) +
    geom_sf(aes(fill = national_alcohol_strategy_written_adopted), colour = "#d0d0d0", size = 0.08) +
    scale_fill_grey(start = 0.9, end = 0.4, na.value = "#f7f7f7") +
    labs(title = "National Alcohol Strategy (Written & Adopted)", subtitle = "Indicator: national_alcohol_strategy_written_adopted", fill = "National Strategy") +
    coord_sf(expand = FALSE) +
    theme_void() +
    theme(plot.title = element_text(size = 18, face = "bold", margin = margin(b = 6)),
          plot.subtitle = element_text(size = 11, margin = margin(b = 8)),
          legend.position = "right")

  print(p_strategy)

```

```{r plot-illegal-alcohol-legislation, fig.height=7, fig.width=11}

  p_illegal <- ggplot(map_data_no_antar) +
    geom_sf(aes(fill = primary_legislation_illegal_alcohol), colour = "#d0d0d0", size = 0.08) +
    scale_fill_grey(start = 0.9, end = 0.4, na.value = "#f7f7f7") +
    labs(title = "Primary Legislation for Illegal Alcohol", subtitle = "Indicator: primary_legislation_illegal_alcohol", fill = "Illegal Alcohol Legislation") +
    coord_sf(expand = FALSE) +
    theme_void() +
    theme(plot.title = element_text(size = 18, face = "bold", margin = margin(b = 6)),
          plot.subtitle = element_text(size = 11, margin = margin(b = 8)),
          legend.position = "right")

  print(p_illegal)

```

## Quick edit: add or edit variables

You can add or edit extra variables simply by editing `map_data` in R and saving a copy if you want to persist changes. Example:

```{r edit-example}
# Example: update in-session metadata manually
map_data$legal_system_family[map_data$name %in% c("singapore")] <- "common law"
map_data$alcohol_policy_status[map_data$name %in% c("singapore")] <- "in place"
```
